---
title: "Project 3: STAT302package Tutorial"
author: Cherry Pan & Yi-Hsuan Wu
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{STAT302package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(STAT302package)
library(tidyverse)
library(devtools)
library(class)
library(palmerpenguins)
library(randomForest)
library(magrittr)
library(dplyr)
library(gapminder)
library(tibble)

```


## 1. Introduction

This is the final project for STAT302. A package called "STAT302package" is created, and it includes four functions: `my_t_test`, `my_lm`, `my_knn_cv`, and `my_rf_cv``.
package collabrator: Cherry Pan & Yi-Hsuan Wu.

The guide to install the package:
```{r, eval = FALSE}
devtools::install_github("Cherry-ty-Pan/STAT302package", build_vignette = TRUE, build_opts = c())
library(STAT302package)
# Use this to view the vignette in the Demo HTML help
help(package = "STAT302package", help_type = "html")
# Use this to view the vignette as an isolated HTML file
utils::browseVignettes(package = "STAT302package")
```


## 2. Tutorial for `my_t.test`

Use the `lifeExp` data from `my_gapminder` for the following tests

### Case I: alternative = "two.sided"
 \begin{align}
  H_0: \mu &= 60,\\
  H_a: \mu &\neq 60.
  \end{align}
  
```{r}
my_t_test(my_gapminder$lifeExp, mu = 60, alternative = "two.sided")
```


### Case II: alternative = "less"
  \begin{align}
  H_0: \mu &= 60,\\
  H_a: \mu &< 60.
  \end{align}
  
```{r}
my_t_test(my_gapminder$lifeExp, mu = 60, alternative = "less")
```


### case III: alternative = "greater"
  \begin{align}
  H_0: \mu &= 60,\\
  H_a: \mu &> 60.
  \end{align}
```{r}
my_t_test(my_gapminder$lifeExp, mu = 60, alternative = "greater")
```


## 3. Tutorial for `my_lm`
```{r}
my_lm(lifeEx ~ gdpPercap + continent, data = my_gapminder)
```

```{r}

```

## 4. Tutorial for `my_knn_cv` using `my_penguins`


## 5. Tutorial for `my_rf_cv`

